---
title: "Oblig 1"
output:
  html_notebook: default
  pdf_document: default
---

## Problem 1


```{r, echo = F}
# Getting data
res_bodyfat <- read.csv("res_bodyfat/res_bodyfat.csv")
attach(res_bodyfat)
```

Aim: Find how well bmi, easily computed as the ratio between weight and squared hight (so mesured in kg/m^2), can be used to predict pbfm, whose measurment involves instead a bioelectrical impedance analysis.

### a

Plotting __bnp__ against __pbfm__, and fitting a simple linear model with a summary.

```{r}
plot(bmi, pbfm)
fit.simpleLinear <- lm(pbfm ~ bmi)
print(summary(fit.simpleLinear))
x = (seq(min(bmi), max(bmi), length = 200))
beta = coef(fit.simpleLinear)
lines(x, beta[1] + beta[2]* x, col = 2, lty = 1, lwd = 2)
```
A linear model seems ok in the center of the bmi. But it is quite off in the lower 1/4 and in the upper 1/4.We can see that the intercept and bmi is strongly significant, indicating a strong correlation between bmi and pdfm. The R-squared is siginificant, but not super high, so the model can probably be improved. To see this clearly we plott indicitative plotts on the fitness.

```{r}
plot(fit.simpleLinear)

```

#### Residuals vs Fitted

Shows the Anscombe plot, wich would idealy present a eaven scattering with no patterns in the points. In the plott we see a clear pattern, and is indicating possiple violation of homoscedasicity. Meaning that \epsiplon must be independent of the index.

#### Normal Q-Q

This plot shows quantile-quantile plot. This plot shows the values of the resiuals in increasing order. To see if the residuals folows a normal distribution. For the assumption of normality to be valid we would expect the values to folow the line in the plot. In the plot a heavy tail and head can be observed. So the assumtion of normality is not strong.

#### Scale-Location

#### Residual vs Leverage

## b

We can see that we dont have any negative number, in such cases it is often benefital to use a logarithmic transformation.

```{r}

fit.log <- lm(pbfm ~ log(bmi))
print(summary(fit.log))
beta = coef(fit.log)

x = log(seq(min(bmi), max(bmi), length = 200))
plot(bmi, pbfm)
lines(exp(x),( beta[1] + beta[2]* x), col = 2) 

```

And to see how good the model is we plot some helping plots

```{r}
plot(fit.log)
```


It was also suggested to use a quadratic model.

```{r}
fit.quad<- lm(pbfm ~ bmi + I(bmi^2))
print(summary(fit.quad))
beta = coef(fit.quad)

x = (seq(min(bmi), max(bmi), length = 200))
plot(bmi, pbfm)
#points(bmi, pbfm)
lines(x, beta[1] + beta[2]*x + beta[3]*x^2, col=2)
```
And the acopening helping plots

```{r}
plot(fit.quad)
```

