---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

# 2.1

## 2.1.1 Basic Concepts

Simple practical problem:

    Identify a relationship that allows us to predict the consumption of fuel, or equivalently, the distance covered per unit of fuel as a function of certain characteristics of a car. We consider data from 203 models of cars in circulation in 1985 in the US, but produced elsewhere. We have 27 different characteristics for each car, four of which are city distance (km/L), engine size (L), number of cylinders, and curb weight (kg). The data is marked for diesel and gasoline.

Some of the data are numerical:

* Quantitative and continuous: City distance, engine size, and curb weight (kg)
* Quantitative and discrete: number of cylinders

Matrix of scaterplots of some variables of car data, stratified by fuel type.

```{r}
auto <- read.table("http://azzalini.stat.unipd.it/Book-DM/auto.dat", header = TRUE)
attach(auto)
#summary(auto)
# Sample size
n = nrow(auto)

# Create dummy variable for fuel: disel = False, gasoline = TRUE
d = fuel == "gas"


# Scater-plot matrix
pairs(auto[ , c("city.distance", "engine.size","n.cylinders","curb.weight")],     
      labels = c("City\ndistance", "Engine\nsize","Number of\ncylinders", "Curb\nweight"),
      col = ifelse(d, 'blue', 'red'), pch = ifelse(d, 1, 2), # 1 is a circle, 2 is a triangle
      cex = 10/sqrt(n))
```

Some are qualitative:

* fuel type: diesel and gasoline

__We will in first phase consider only two explanatory variables:__

* Engine size, fuel type

To study the relationship between quantitative variables we usually make a grafic representation.

```{r}
### figure 2.2 ###
plot(engine.size, city.distance, type = "n", ylab = "City distance (km/L)",
     xlab = "Engine size (L)", xlim = c(1, 5.5))
points(engine.size[d], city.distance[d], col = 4, pch = 1)
points(engine.size[!d], city.distance[!d], col = 2, pch = 2)
legend('topright', pch = c(1, 2), col = c(4, 2),
       legend = c("Gasoline  ","Diesel"))
```

We first suggest a simple linear regression line: $y = \beta_{0} + \beta_{1} x + \epsilon$, where y represents city distance, x fuel type, and $\epsilon$ is a nonobservable random 'error', with expected value 0 and variance $\sigma^2$. For simplicity we consider no correlation between error terms and y.
We need to find a setimator for the unknown variales $\beta_{0}$ and $\beta_{1}$. To do so wee need to user the method of least squares. Wich means the finding the min of the function:

\center $B(\beta)=\Sigma_{i=1}^n \{y_{i} -f(x_{i};\beta) \}^2 = ||y-f(x;\beta)||^2$ \center

The last expressison is showing the matrix notation for representing $y = (y1,...,y_{n})^T; (f(x_{1};\beta),...,f(x_{m};\beta))^T;$.

__But from what we can see in the plot of 'city distance against engine size', a linear model might not be the best.__
Therefor we might consider a polynomial form, since it is easy and quick. The polynom will look something like:
\center $f(x;\beta) = \beta_{0} + \beta_{1} x + ... + \beta_{p-1} x^{p-1}$ \center

Because of the simplicity if a polynomial we can wirte f in matrix form with low effort,
\center $f(x;\beta)= X\beta$ \center
where x is an n x p matrix, called the design matrix, defined by
\cnter $X=(1,x,...,x^{p-1})$ \center
where x is the vector of the observations of the explanatory variable, and the varius columns of X contains powers of order from 0 to p-1 of elements of x.

On the matrix form the solution to the minimaztion problem is
\center $\hat{\beta} = (X^{T}X)^{-1}X^{T}y$



### 2.1.3 Multivariable Responses

In many cases we would like more than one responsvariable. In the car example we for example would like higway and city distance to be out responsvariables. 

If we create q response variables, we can construct a matrix Y such that the columns that contain the q variables. In pur example we have q = 2, and then 

\center $Y = ((CityDistance), (HighwayDistance))$. \center

If we create q models of linear regression, each of the type seen when we only had one response, using the same regression matrix X for each, we obtain

\center $Y = XB + E$ \center \hfill    (2.20)

where B us the matrix formed of __q__ columns of dimention __p__, each representing the regression parameters for the corresponding column of Y, and matrix R is made up of error terms. Here aswell each of its columns reffers to the corresponding column of Y, with condition that

\center $var[\tilde{E}_{i}] = \Sigma$ \center

where $\tilde{E}_{i}^{T}$ represent the i'th row E, for i = 1, ... , n, and $\Sigma$ is a variance matrix of dimention q x q independent of i, wich expresses the correlation structure between the error components and therefor also between the response variables. Equation (2.20) constitutes a model of _'multivariate multiple linear regression'_, where 'multivariate' refers to the fact that we have multiple (q) response variables, and 'multiple' means we have multiple (p) explanetory varialbes.

The natural extention of the least square criterion will be similar to the one not multivariate. But with q response variables. This gives the model

\center $\hat{B} = (X^{T}X)^{-1} x^{T} Y$ \center \hfill (2.21)

which is just the juxtaposition of q vectors estimated for each response variable.


## Likelihood

### 2.3.1 General Concepts





