---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

# 2.1

## 2.1.1 Basic Concepts

Simple practical problem:

    Identify a relationship that allows us to predict the consumption of fuel, or equivalently, the distance covered per unit of fuel as a function of certain characteristics of a car. We consider data from 203 models of cars in circulation in 1985 in the US, but produced elsewhere. We have 27 different characteristics for each car, four of which are city distance (km/L), engine size (L), number of cylinders, and curb weight (kg). The data is marked for diesel and gasoline.

Some of the data are numerical:

* Quantitative and continuous: City distance, engine size, and curb weight (kg)
* Quantitative and discrete: number of cylinders

Matrix of scaterplots of some variables of car data, stratified by fuel type.

```{r}
auto <- read.table("http://azzalini.stat.unipd.it/Book-DM/auto.dat", header = TRUE)
attach(auto)
#summary(auto)
# Sample size
n = nrow(auto)

# Create dummy variable for fuel: disel = False, gasoline = TRUE
d = fuel == "gas"


# Scater-plot matrix
pairs(auto[ , c("city.distance", "engine.size","n.cylinders","curb.weight")],     
      labels = c("City\ndistance", "Engine\nsize","Number of\ncylinders", "Curb\nweight"),
      col = ifelse(d, 'blue', 'red'), pch = ifelse(d, 1, 2), # 1 is a circle, 2 is a triangle
      cex = 10/sqrt(n))
```

Some are qualitative:

* fuel type: diesel and gasoline

__We will in first phase consider only two explanatory variables:__

* Engine size, fuel type

To study the relationship between quantitative variables we usually make a grafic representation.

```{r}
### figure 2.2 ###
plot(engine.size, city.distance, type = "n", ylab = "City distance (km/L)",
     xlab = "Engine size (L)", xlim = c(1, 5.5))
points(engine.size[d], city.distance[d], col = 4, pch = 1)
points(engine.size[!d], city.distance[!d], col = 2, pch = 2)
legend('topright', pch = c(1, 2), col = c(4, 2),
       legend = c("Gasoline  ","Diesel"))
```

We first suggest a simple linear regression line: $y = \beta_{0} + \beta_{1} x + \epsilon$, where y represents city distance, x fuel type, and $\epsilon$ is a nonobservable random 'error', with expected value 0 and variance $\sigma^2$. For simplicity we consider no correlation between error terms and y.
We need to find a setimator for the unknown variales $\beta_{0}$ and $\beta_{1}$. To do so wee need to user the method of least squares. Wich means the finding the min of the function:

\center $B(\beta)=\Sigma_{i=1}^n \{y_{i} -f(x_{i};\beta) \}^2 = ||y-f(x;\beta)||^2$ \center

The last expressison is showing the matrix notation for representing $y = (y1,...,y_{n})^T; (f(x_{1};\beta),...,f(x_{m};\beta))^T;$.

__But from what we can see in the plot of 'city distance against engine size', a linear model might not be the best.__



